Naming Anonymous JavaScript Functions

### 2011-04-11 trunk/papers/splash-wavefront2011/sigplanconf-template.tex
notes from our IRC verbatim

So what do JavaScript developers do in practice?  To get empirical evidence we analyzed the source code of ten well-known JavaScript projects. For every project, the total number of functions and the number of functions with an identifier shown in Table~\ref{js-functions}. The average ratio of named functions to all functions is less then 7 percent and, excepting one project {\textit Prototype}, the ratio does not exceed 13 percent. Among all functions only a very limited number of them (116 functions) are defined by the {\small\texttt{Function}} constructor. Our analysis does not include the functions defined dynamically by {\small\texttt{eval}} function or {\small\texttt{new Function()}}; these cases would only make the ratio even smaller.  Therefore, we conclude that a large proportion of JavaScript functions are anonymous.

\begin{table}
\centering
\scalebox{0.8}{
  \begin{tabular}{ | l | l | l | l | l |}
  \hline
   Project & Description & Total & Named \\ 
  \hline 
   Closure & Google Web Library & 9195 & 208(2\%) \\ 
  \hline 
   DoJo & JavaScript Toolkit & 18676 & 2810(15\%) \\ 
  \hline 
   ExtJS & JavaScript Framework & 37717 & 1184(3\%) \\ 
  \hline 
   Firebug & Web Development Tool & 3424 & 406(11\%) \\ 
  \hline 
   jQuery & JavaScript Library & 422 & 23(5\%) \\ 
  \hline 
   MochiKit &  JavaScript Library & 1866 & 37(1\%) \\ 
  \hline 
   MooTools & JavaScript Framework & 625 & 7(1\%) \\ 
  \hline 
   Prototype & JavaScript Framework & 645 & 203(31\%) \\ 
  \hline 
   Scriptaculous &  JavaScript Library & 1092 & 208(19\%) \\ 
  \hline 
   YUI &  Yahoo UI Library & 22346 & 922(4\%) \\ 
  \hline 
   All &  & 96008 & 6008(6.3\%) \\ 
  \hline 
  \end{tabular}
  }
\caption{The total number of functions and the number of named functions (and percent named)  in ten large JavaScript projects. See appendix 1 for the project citations.}
\label{js-functions} 
\end{table}

@@

We parse the JavaScript and search the resulting syntax tree for function body nodes. For each function body, we apply the algorithm outlined in Algorithm~\ref{alg1}). The basic idea is to look for parent nodes with helpful identifiers as outlined in  Table~\ref{table:node-types}. In simple cases the parent node will have a simple identifier; otherwise we construct an identifier from the expression as described in \ref{sec:general-element-naming}.  At each step we track how the parent and child are related.  If the parent expression is an array or object literal then the current node value is \textit{partOf} the parent node value. Otherwise the node expression value is \textit{theSame} as or different \textit{contributesTo} parent value. The algorithm ends when we reach a node which is an assignment or a statement which does not return any value (i.e., the node is not an expression).  The algorithm result is an \textit{object consumption summary}, a list of collected data at every visited parent node of the function body.

Next we rank the collected identifiers and strings to elide less relevant data from the summary. Finally we concatenate the items with marker symbols to construct a string from the consumption summary. The result contains identifiers (i.e., variable and property names) and strings available in the source code plus some explanatory tokens. It compactly explains the function object creation, consumption and assignment.

>>>>

So what do JavaScript developers do in practice?  To get empirical evidence we analyzed the source code of ten well-known JavaScript projects \footnote[1]{We did not perform any preprocess to exclude third-party or repeated files in the provided source code bundles.}. For every project, the total number of functions and the number of functions with an identifier shown in Table~\ref{js-functions}. The average ratio of named functions to all functions is less then 7 percent and, excepting one project {\textit Prototype}, the ratio does not exceed 13 percent. Among all functions only a very limited number of them (116 functions) are defined by the {\small\texttt{Function}} constructor. Our analysis does not include the functions defined dynamically by {\small\texttt{eval}} function or {\small\texttt{new Function()}}; these cases would only make the ratio even smaller.  Therefore, we conclude that a large proportion of JavaScript functions are anonymous.

@@

We parse the JavaScript and search the resulting syntax tree for function body nodes. For each function body, we apply the algorithm outlined in Algorithm ~\ref{alg1}. The basic idea is to look for parent nodes with helpful identifiers as outlined in  Table~\ref{table:node-types}. In simple cases the parent node will have a simple identifier; otherwise we construct an identifier from the expression as described in \ref{sec:general-element-naming}.  At each step we track how the parent and child are related.  If the parent expression is an array or object literal then the current node value is \textit{partOf} the parent node value. Otherwise the node expression value is \textit{theSame} as or different \textit{contributesTo} parent value. The algorithm ends when we reach a node which is an assignment or a statement which does not return any value (i.e., the node is not an expression).  The algorithm result is an \textit{object consumption summary}, a list of collected data at every visited parent node of the function body.

Next we rank the collected identifiers and strings to elide less relevant data from the summary. 
\begin{verbatim}
	salmir	if during the walking up we reach a "return" node and the outer function is a "closure function" we jump to the parent node of "closure function"
	johnjbarton	inside the while correct?
	salmir	right.
	salmir	there is also another effect, but not in the algorithm. The parent context name will become the closure functions parent context name.
johnjbarton	can we handle this case with the end of the while loop:
	johnjbarton	n = n.parent
	salmir	yes.
	johnjbarton	to
	johnjbarton	n = getNextNode(n)
	salmir	getParentNode();
	johnjbarton	then explain getNextNode is either n.parent or the funky other thing

		so now we are done with the summary and we process it with another algorithm
	salmir	yes.drop case 2 unless we have at least one of 1 or 3?
	salmir	yes.
	johnjbarton	why?because the name from the function call is less valuable than the target
	salmir	because those identifiers give some information about the function object directly, but the function calls are indirect.
johnjbarton	ok we need an example for both paths
salmir	line 13
	salmir	we drop op()
	salmir	for the other case when we keep function call, you can use figure 4.
	salmir	jquery example.
	johnjbarton	which one, inner or outer
	salmir	inner
	johnjbarton	I don't see it
	salmir	right, outer.

salmir	if there are more than on function call we just use the closes one to the function defintion.
	salmir	foo(bar(baz("MyFun", function(){})) -> baz[MyFun]()
	johnjbarton	so if we have two case 2 and zero 1 or 3, then pick the early case 2?
	salmir	yes.

johnjbarton	so the next step is to concatenate
	johnjbarton	and you just remove dups or?
	salmir	just that.
	johnjbarton	and the reason for << is no identifier is found
	salmir	yes, or internal nodes.
	johnjbarton	what is an internal node?>
	salmir	foo(bar(baz((function(){})))
	salmir	<<<
	salmir	myObject = foo(bar(baz((function(){})))
        salmir	here foo, bar baz all are internal nodes.
	salmir	so we just get myObject< as name
\end{verbatim}
Finally we concatenate the items with marker symbols to construct a string from the consumption summary. The result contains identifiers (i.e., variable and property names) and strings available in the source code plus some explanatory tokens. It compactly explains the function object creation, consumption and assignment.
